% #############################################################################
% Acknowledgments
% !TEX root = main.tex
% #############################################################################

This work was partially supported by national funds through \acl{FCT} (\acs{FCT}) and \acl{IST} (\acs{IST}) within several funding sources.
First, the work was supported through the UID/EEA/50009/2013 project.
Second, the work was supported alongside the LARSyS - \acs{FCT} Project 2022.04485.PTDC hosted by \acs{IST}.
At last, the work was supported by both BL89/2017-\acs{IST}-ID and PD/BD/150629/2020 grants from my \acl{PhD} (\acs{PhD}) funding program.

I would like to express my sincere gratitude to the following individuals for their contributions to my research and for making my journey as a Doctoral Researcher an enjoyable and enriching experience.
First and foremost, I would like to thank my supervisors, Prof. Jacinto C. Nascimento and Prof. Nuno J. Nunes, for their guidance, support, and encouragement throughout my doctoral studies.
I would like to acknowledge the support of my family, friends, and colleagues, who have provided me with unwavering support and encouragement throughout my academic journey.

There are several \acl{MSc} (\acs{MSc}) students who I informally co-supervised or provided mentorship that I would like to acknowledge, as well.
I would like to thank the following \acs{MSc} students, whom I had the pleasure of informally co-supervising: Hugo Lencastre, N\'{a}dia Mour\~{a}o, Jo\~{a}o Fernandes, Miguel Bastos, and Vicente Sobral.
Your hard work and dedication to our research projects were greatly appreciated, and I am proud of the work we accomplished together.
Additionally, I would like to thank the \acs{MSc} students whom I provided mentorship to: Jo\~{a}o Bernardo, Madalena Pedreira, Pedro Diogo, Margarida Morais, and Mauro Machado.
It was a privilege to work with each of you, and I am confident that you will continue to achieve great things in your future endeavors.

Additionally, I am grateful for the invaluable assistance provided by my colleagues from the \acl{HCII} (\acs{HCII}) at \acl{CMU} (\acs{CMU}), during my Visiting Scholar program in the United States.
Specifically, I would like to extend my heartfelt gratitude to Prof. John Zimmerman.
The invaluable feedback and discussions across this research have played a crucial role in its success.
Similarly, I would like to thank Prof. Adam Perer from the \href{https://dig.cmu.edu/}{DIG Lab} for the opportunity of being part of a great team.
Thanks for the collaboration and the hard work of both.
I would like to extend my sincerest gratitude to Venkatesh Sivaraman, Katelyn Morrison, Angel (Alex) Cabrera from the \acs{HCII} @ \acs{CMU}, and Nari Johnson from the \acs{ML} @ \acs{CMU}, for their valuable insights and suggestions.

I am indebted to those who gave their time and expertise to evaluate our work.
Not forgetting the great support of Prof. Dominik Moritz during this path.
I would like to extend my heartfelt thanks to the following individuals for their contributions to this work, such as Frank Elavsky, Conrad Borchers, Franklin (Mingzhe) Li, Neeta Khanuja, William Epperson.
I would also like to acknowledge the support of these \acs{PhD} students, who gave me significant insights to this research work.
Their dedication and comments have not gone unnoticed and are greatly appreciated.
Thank you to the following faculty members, such as Jeffrey Bigham, Patrick Carrignton, Ken Holstein, Brad Mayers, Ken Koedinger, and Vicent Aleven for their opinions and appreciations of the work.
Last but not least, thank you Qianou (Christina) Ma, Yunzhi Li, and Sanika Moharana for the companionship.

Several clinicians and clinical institutions supported and provided a great collaboration to this \acs{PhD} thesis.
Because of that, I would like to thank Dr. Clara Aleluia and her radiology team at \acl{HFF} (\acs{HFF}) for valuable insights and for helping to use the assistant on a daily basis.
\acs{HFF} is an \acl{EPE} (\acs{EPE}) clinical institution in Portugal.
Equally important, this research involves collaboration between two central public cancer centers in Portugal, namely the \acl{IPO} (\acs{IPO}) institutions located in Lisbon and Coimbra.
From \acl{IPOL} (\acs{IPOL}), I would like to thank the medical imaging teams of Dr. Jos\'{e} Carlos Marques and Dr. Jos\'{e} Ven\^{a}ncio.
From \acl{IPOC} (\acs{IPOC}), we would like to thank the radiology department director and all team of Dr. Id\'{i}lio Gomes.
Moreover, I would like to provide my acknowledgments to Dr. Em\'{i}lia Vieira and Dr. C\'{a}tia Pedro from \acl{HSM} (\acs{HSM}), a clinical institution of \acl{CHULN} (\acs{CHULN}).
For participation, as well as from all other clinical institutions, I also want to thank all clinicians from the radiology department of \acl{HB} (\acs{HB}), a clinical institution of \acl{CHBM} (\acs{CHBM}).
From \acl{SAMS} (\acs{SAMS}), a great thank to Dr. Cristina Ribeiro da Fonseca who, among others, is giving us crucial information for the {\it BreastScreening} research work.
From \acl{CHLO} (\acs{CHLO}), I would like to convey Dr. Zacharoula Sidiropoulou and Dr. Tiago Saldanha for their participation, as well as for networking with their medical teams.
Next, I would like to acknowledge the collaboration with \acl{CHTMAD} (\acs{CHTMAD}) and \acl{JCCC} (\acs{JCCC}) thanks to Dr. Jo\~{a}o Maria Abrantes and his team.
In addition, it is important also to acknowledge the efforts made by Dr. Duarte Freitas from \acl{MMC} (\acs{MMC}), as well as Dr. Mariana Correia from \acl{CF} (\acs{CF}) and Dr. Zita Seabra from \acl{HVFX} (\acs{HVFX}), who among others provided excellent feedback for our work.
Last, thanks to Dr. Willian Schmitt from \acl{CHUSA} (\acs{CHUSA}).
I want to thank all clinicians and clinical institutions who generously contributed their time and expertise in assessing our research efforts.